---
title: "Analysis"
author:
- "Haojia Li"
- "Zihan Yang"
- "Tianran Zhang"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    code_folding: hide
    theme: readable
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width=6, fig.height=4, fig.align = "center")
```

```{r load data}
# Load Data
# load("/Users/zihanyang/Documents/Capstone_ESP/Data/RData/ESP_reg.RData")
load("../data/ESP_reg.RData")
```


```{r packages}
# Load package
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(kableExtra)
library(knitr)
library(arsenal)
```

```{r boxplot.test function}
# produce all the pairs within a categorical vector
produce.full.comp <- function(x) {
  x <- as.factor(x)
  n <- nlevels(x)
  l <- levels(x)
  comp <- list()
  k <- 0
  for(i in 1:(n-1)) {
    for(j in (i+1):n) {
      k <- k+1
      comp[[k]] <- l[c(i,j)]
    }
  }
  return(comp)
}



# draw the distribution of numeracy score by covariates (x)
boxplot.test <- function(x, y = "numeracy", data = ESP.reg, 
                         pairwise = T, filter.sig = T, comp = NULL,
                         title = NULL, xlab = "", ylab = "", label.y = 25) {
  if(is.null(title))
    title <- paste0("Numeracy distribution by ", x)
  comp.pairs <- produce.full.comp(unlist(data[x]))
  comp.result <- compare_means(as.formula(paste0(y,"~",x)), data = data)
  if(is.null(comp)) {
    if(filter.sig)
      comp <- comp.pairs[which(comp.result$p.signif!="ns")]
    else
      comp <- comp
  }
  p <- ggboxplot(data=data, x=x, y=y, color=x, palette = "jco", legend="",
                 title = title, xlab=xlab, ylab=ylab) +
    stat_compare_means(aes(label =ifelse(as.numeric(..p.format..)<0.001,
                                         paste0(..method.., ", p<0.001"),
                                         paste0(..method.., ", p=", round(as.numeric(..p.format..),3)))),
                       label.y = label.y)
  if(pairwise)
    p <- p + stat_compare_means(aes(label = ..p.signif..),
                                symnum.args = list(cutpoints = c(0, 0.001, 0.01, 0.05, 1), 
                                                   symbols = c("***", "**", "*", "ns")),
                                comparisons = comp)
  return(p)
}


```


# ESP

## Differential  Analysis


### Summary 

1. The distributions of numeracy scores are significant different in the subgroups of `race` , `hisp` , `education` , `income` , `age.cat`.

2. No matter using  not not-poor/poor (corr=0.136, p-value < 0.001) or good/not good(corr=0.175, p-value < 0.001)as the self-reported health status outcome variable, *numeracy score* and self-reported status have positive significant corrrealtion. 

3. No matter using  not poor/poor or good/not good as the self-reported health status outcome variable, `education` (chisq p-value < 0.001) and `income` (chisq p-value < 0.001) subgroups, and `numeracy score` have the significant effects at the  two-sided significance level of 0.05.

### Numeracy 

#### Numaracy vs. Binary Vriable
When the ordinal variable has only two levels, wilcoxon test will be performed directly.

##### Numaracy vs. Health Status

`Differential Analysis` <br>
-  numerace vs. health status: good/not good <br>
There is enough evidence to reject the null hypothesis that the distribution of numeracy scores are the same is good/not good health status subgroup. (p-value < 0.001)

```{r numeracy by good health or not}
boxplot.test("health_status_good", pairwise = F, label.y = 20, 
             title = "Numeracy distribution by good health or not")
```

-  numeracy vs. health status: not poor/poor <br>
There is enough evidence to reject the null hypothesis that the distribution of numeracy scores are the same is not poor/poor health status subgroup. (p-value < 0.001)
```{r numeracy by poor health or not}
boxplot.test("health_status_poor", pairwise = F, label.y = 20, 
             title = "Numeracy distribution by poor health or not")
```

`Correlation` <br>
[Point-Biserial Correlation](https://en.wikipedia.org/wiki/Point-biserial_correlation_coefficient): 1 continuous variable (numeracy score) and 1 binary variable (health status) <br>

No matter using  not not-poor/poor (corr=0.136, p-value < 0.001) or good/not good  (corr=0.175, p-value < 0.001)as the self-reported health status outcome variable, numeracy and self-reported status have positive significant corrrealtion. 

```{r correlation}
# Poor/Not Poor
# wilcox
wilcox <- wilcox.test(numeracy~health_status_poor, data=ESP.reg)
ap1 <- round(wilcox$p.value,8)
# correaltion: Point-Biserial Correlation: 1 continuous variable and 1 binary variable
ESP.reg$health_num_poor <- ifelse(ESP.reg$health_status_poor=="Not Poor",1,0)
health_num_poor<- with(ESP.reg,cor.test(numeracy,health_num_poor,alternative = "two.sided", method="pearson",conf.level = 0.95,na.action="na.omit"))
ap2 <-round(health_num_poor$estimate,8)
names(ap2) <- NULL
ap3 <-round(health_num_poor$p.value,8)

# Good/Not Good
# wilcox
wilcox <- wilcox.test(numeracy~health_status_good, data=ESP.reg)
ag1 <-round(wilcox$p.value,8)
# correaltion: Point-Biserial Correlation: 1 continuous variable and 1 binary variable
ESP.reg$health_num_good <- ifelse(ESP.reg$health_status_good=="Good",1,0)
health_num_good<- with(ESP.reg,cor.test(numeracy, health_num_good,alternative = "two.sided", method="pearson",conf.level = 0.95,na.action="na.omit"))
ag2 <-round(health_num_good$estimate,8)
names(ag2) <- NULL
ag3 <-round(health_num_good$p.value,8)


df <-data.frame(`Correlation` = c("Numeracy vs. Poor/Not Poor", "Numeracy vs. Good/Not Good"),`Point-Biserial Correlation` = c(round(ap2,5),round(ag2,5)), `P.vale` = c("<0.001", "<0.001"))
kable(df)%>%
  kable_styling("striped", full_width = FALSE)
```

-  numeracy vs. hisp <br>
There is enough evidence to reject the null hypothesis that the distribution of numeracy scores are the same is hisp/non-hisp subgroup.(p-value < 0.001)

```{r numeracy by ethnicity}
boxplot.test("hisp", pairwise = F, label.y = 20, 
             title = "Numeracy distribution by ethnicity")
```

-  numeracy vs. gender <br>
There is enough evidence to reject the null hypothesis that the distribution of numeracy scores are the same is poor/not poor health status subgroup. (p-value < 0.001)
```{r numeracy by gender}
boxplot.test("gender", pairwise = F, label.y = 20)
```

#### Numaracy vs. Multi-level Vriable
When the ordinal variable has more than two levels,kruskal-wallis test first be performed, pairwise wilcoxon tests will be performed as follows . <br>
There is enough evidence to reject the null hypothesis that the distribution of numeracy scores are the same in income subgroups. (p-value < 0.001). Specifically, further comparison indicating that the distribution of numeracy scores are different in income level pairwisely.(all pairs with p-value < 0.001 )

```{r numeracy by income}
boxplot.test("income.cat", title = "Numeracy distribution by house hold income")
```

There is enough evidence to reject the null hypothesis that the distribution of numeracy scores are the same is education subgroup. (p-value < 0.001) Specifically, further comparison indicating that the distribution of numeracy scores are different in education level pairwisely.(all pairs with p-value < 0.001)


```{r}
boxplot.test("educ", title = "Numeracy distribution by education level") +
  rotate_x_text(6) +
  font("x.text", size = 8)
```

There is enough evidence to reject the null hypothesis that the distribution of numeracy scores are the same in age subgroup. (p-value = 0.02) Specifically, further pairwise comparison indicating that the distribution of numeracy scores are significantly different (age 21-44 vs. age 45-64)(p-value = 0.0069), and a(ge 21-44 vs. age 65+) (p-value = 0.023)

```{r}
boxplot.test("age.cat", title = "Numeracy distribution by age level")
```

There is enough evidence to reject the null hypothesis that the distribution of numeracy scores are the same in race subgroup. (p-value < 0.001) Specifically, further pairwise comparison indicating that the distribution of numeracy scores are significantly different in white vs. black (p-value < 0.001), and white vs.other (p-value <0.001)

```{r}
boxplot.test("race")
```


### Health Status 
Because all the variables intended to put in the regression are categorical except the numeracy. We would use chisq test to check self-reported health status across different subgroup(gender, age, education, income, race, hisp) seperately. 

1. Health status: good/not good <br>
There is enough evidence to reject the null hypothesis that the proportions of good self-reported health status are the same among different education subgroups at the two-sided significance level of 0.05.(p-value < 0.001) And there is enough evidence to reject the null hypothesis that the proportions of good self-reported health status are the same among different income subgroups at the two-sided significance level of 0.05.  (p-value < 0.001)

```{r}
cat_health_good <- tableby(health_status_good ~ gender + age.cat + educ + income.cat + race +hisp, cat.test="chisq", data=ESP.reg)
cat_health_good <- as.data.frame(tests(cat_health_good))

pval_good <- t(cat_health_good$p.value) %>%
  as.data.frame()

pval.format <- function(p) {
  return(ifelse(p < 0.001, "<0.001", round(p,3)))
}

pval.symbol <- function(p) {
  return(case_when(
    p < 0.001 ~ "***",
    p < 0.01 ~ "**",
    p < 0.05 ~ "\\*",
    TRUE ~ ""
  ))
}
pval_good <- as.data.frame(rbind(pval.format(cat_health_good$p.value),
                                 pval.symbol(cat_health_good$p.value)))
colnames(pval_good) <- c("Gender", "Age", "Education", "Income", "Race", "Ethnicity")
rownames(pval_good) <- c("P-value", "Significance")
kable(pval_good, caption = "Pearson's Chi-squared Test between Good or Not Good Health Status and Demographical Covarites") %>%
  kable_styling()

```

```{r include=FALSE}
colnames(pval_good) <- c("Gender", "Age", "Education", "Income", "Race", "Ethnicity")
rownames(pval_good) <- c("P-value", "Significance")
kable(pval_good, caption = "Pearson's Chi-squared Test between Good or Not Good Health Status and Demographical Covarites") %>%
  kable_styling()

cat_health_good<- cat_health_good%>%mutate(p.value = ifelse(p.value < 0.001, "<0.001", round(p.value,3)))
kb_good <- kable(cat_health_good)
kb_good%>%kable_styling("striped", full_width = FALSE)

```


2. Health status: not poor/poor <br>
There is enough evidence to reject the null hypothesis that the proportions of not-poor self-reported health status are the same among different education subgroups at the two-sided significance level of 0.05.(p-value < 0.001) And there is  enough evidence to reject the null hypothesis that the proportions of not-poor self-reported health status are the same among different income subgroups at the two-sided significance level of 0.05. (p-value < 0.001)

```{r}
cat_health_poor <- tableby(health_status_poor ~ gender + age.cat + educ + income.cat + race +hisp, cat.test="chisq", data=ESP.reg)
cat_health_poor <-as.data.frame(tests(cat_health_poor))

pval_good <- as.data.frame(rbind(pval.format(cat_health_poor$p.value),
                                 pval.symbol(cat_health_poor$p.value)))
colnames(pval_good) <- c("Gender", "Age", "Education", "Income", "Race", "Ethnicity")
rownames(pval_good) <- c("P-value", "Significance")
kable(pval_good, caption = "Pearson's Chi-squared Test between Not Poor or Poor Health Status and Demographical Covarites") %>%
  kable_styling()

```

```{r include=F}
cat_health_poor <- cat_health_poor%>%mutate(p.value = ifelse(p.value < 0.001, "<0.001", round(p.value,3)))
kb_poor <- kable(cat_health_poor)
kb_poor%>%kable_styling("striped", full_width = FALSE)
```








