---
title: "piaac"
author: "Tianran Zhang"
date: "6/23/2020"
output:
  html_document:
    code_folding: hide
    theme: readable
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(knitr)
library(kableExtra)
# data manipulation
library(tidyverse)

# Draw ROC plot
library(ROCR)
library(pROC)
library(simr)

load("../Data/PIAAC_reg.RData")
```

```{r}
rocplot <- function(truth, pred, ...){
  predob = prediction(pred, truth)
  perf = performance(predob, "tpr", "fpr")
  plot(perf, colorize=F, ...) 
  area = auc(truth, pred)
  area = format(round(area, 4), nsmall = 4)
  text(x = .8, y = .1, labels = paste("AUC = ", area))
  segments(x0 = 0, y0 = 0, x1 = 1, y1 = 1, col = "gray", lty = 2)
}

```


## PIAAC  
Health status (Binary):  Good VS. Not Good      
* Predictor: numeracy1 score        
* Covariates: education, income, race, ethincity   
* Additional predictors: literacy1, problem solving    

```{r}
n1 <- nrow(PIAAC.reg)
set.seed(135)
id1 <- sample(n1, n1 * .6)
PIAAC.training <- PIAAC.reg[id1, ]
PIAAC.test <- PIAAC.reg[-id1, ]

PIAAC.reg
```

## Consider numeracy1 as the only predictor    


```{r, warning=FALSE}
varlist2 <- c("numeracy1", "educ", "income.cat", "race", "hisp",
              "literacy1", "problem_solving1")

full7 <- glm(health_status_good ~ numeracy1 + educ + income.cat +
               race + hisp, data = PIAAC.reg, family = "binomial")
final7 <- step(full7, direction = "both", trace = 0)
# summary(final7)
fit7 <- glm(health_status_good ~ numeracy1 + educ + income.cat, 
            data = PIAAC.reg, family = "binomial")

multi_summary7 <- exp(cbind(coef(final7), confint(final7))[-1, ]) %>% as.data.frame()
colnames(multi_summary7)[1] <- "OR"
multi_summary7$sig <- ifelse(summary(final7)$coefficients[-1, 4] < 0.05, "*", "")

# power calculation
power7 <- powerSim(final7, nsim = 100, progress = FALSE)
auc7 <- auc(PIAAC.reg$health_status_good, predict.glm(final7, PIAAC.reg, type = "response", na.action = na.omit))
kable(multi_summary7) %>%
  kable_styling()

# Calculate error rate
fit7.train <- glm(health_status_good ~ numeracy1 + educ + income.cat, 
            data = PIAAC.training, family = "binomial")
pred7 <- predict(fit7.train, PIAAC.test, type = "response")
pred7 <- ifelse(pred7 > .5, "Not Good", "Good")
err7 <- mean(pred7 != PIAAC.test$health_status_good)
```

Step-wise multivariate regression model was conducted on PIAAC data. The final model has three predictors :numeracy1, education and income. The AUC is `r auc7[1]` with a power of `r power7$x/power7$n`. The prediction error rate is `r round(err7, 3)`.    

```{r, message=FALSE}
# ROC curve
rocplot(PIAAC.reg$health_status_good, predict.glm(final7, PIAAC.reg, type = "response", na.action = na.omit), title = "ROC plot")
title(main = "ROC plot")
```


## Consider two additional predictors: literacy1 and Problem solving abilities     
```{r, warning=FALSE}
full8 <- glm(health_status_good ~ numeracy1 + educ + income.cat +
               race + hisp + literacy1 + problem_solving1, 
             data = PIAAC.reg, family = "binomial")
final8 <- step(full8, direction = "both", trace = 0)
# summary(final8)
fit8 <- glm(health_status_good ~ numeracy1 + educ + income.cat +
               literacy1 + problem_solving1, data = PIAAC.reg, 
            family = "binomial")
# summary(final8)
multi_summary8 <- exp(cbind(coef(final8), confint(final8))[-1, ]) %>%
  as.data.frame()
colnames(multi_summary8)[1] <- "OR"
multi_summary8$sig <- ifelse(summary(final8)$coefficients[-1, 4] < 0.05, "*", "")

# power calculation
power8 <- powerSim(final8, nsim = 100, progress = FALSE)
auc8 <- auc(PIAAC.reg$health_status_good, predict.glm(final8, PIAAC.reg, type = "response", na.action = na.omit))
kable(multi_summary8) %>%
  kable_styling()

# Calculate error rate
fit8.train <- glm(health_status_good ~ numeracy1 + educ + income.cat +
               problem_solving1, data = PIAAC.training, 
            family = "binomial")
pred8 <- predict(fit8.train, PIAAC.test, type = "response")
pred8 <- ifelse(pred8 > .5, "Not Good", "Good")
err8 <- mean(pred8 != PIAAC.test$health_status_good)
```

Step-wise multivariate regression model was conducted. The final model has four predictors: numeracy1, educ, income and problem solving. The AUC is `r auc8[1]` with a power for numeracy1 of `r power8$x/power8$n`. The prediction error rate is `r round(err8, 3)`. 

```{r}
rocplot(PIAAC.reg$health_status_good, predict.glm(final8, PIAAC.reg, type = "response", na.action = na.omit), title = "ROC plot")
title(main = "ROC plot")
```



# Summary 

## PIAAC
```{r}
PIAAC_h1 <- c("health status: Good vs Not Good", round(auc7[1], 3), 
            power7$x/power7$n, round(err7, 4), "\\*","\\*", "\\*","", "","/", "/")
PIAAC_h2 <- c("health status: Good vs Not Good", round(auc8[1], 3), 
            power8$x/power8$n, round(err8, 4), "\\*","\\*", "\\*","", "","\\*", "\\*")

PIAAC.result <- data.frame(PIAAC_h1, PIAAC_h2) %>% 
  t() %>% 
  as.data.frame()
rownames(PIAAC.result) <- NULL
colnames(PIAAC.result) <- c("Outcome", "AUC", "Power", "Prediction Error Rate",
                        "numeracy", "educ", "income", "race",
                        "hisp", "problem solving", "literacy")
PIAAC.result %>%
  kable(caption = "PIAAC's multivariates regression summary") %>%
  kable_styling() %>%
  add_header_above(c("Response" = 1, "Performance"=3,"Predictor"=1,"Covariates"=4, "Additional Predictor" = 2))
```


