---
title: "PICCA"
author: "Tianran Zhang"
date: "5/24/2020"
output:
  html_document:
    code_folding: hide
    toc: yes
    theme: readable
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# Load data
library(haven)

# EDA
library(tidyverse)
library(table1)
library(knitr)

# Draw ROC plot
library(ROCR)
library(pROC)

# knit tables
library(knitr)
library(kableExtra)

# Force R not to use exponential notation (e^(-2))
options("digits"=3) # maintain 3 digits
```


```{r results='hide'}
load("../Data/PIAAC.RData")
PIAAC
str(PIAAC)
sum(is.na(PIAAC))
```

# Baseline analysis
Descriptive table
```{r}
# Generate table1
table1(~ health_score + health_status + health_status_1 + health_status_2 + numeracy + 
         educ + income.cat + race + literacy + problem_solving, data = PIAAC)
```


# Regression Analysis 

```{r}
PIAAC_clean <- PIAAC %>%
  filter(health_status != "Unknown") %>%
  mutate(health_status_1 = factor(health_status_1),
         health_status_2 = factor(health_status_2))
```


## Univariate ordinal logistic regression
**With health status 1 as outcome**    
health_status_1:     
  Poor: health score >=5;    
  Not poor: health score < 5.   
  
```{r}
univ <- function(var){
  form = formula(paste0("health_status_1 ~ ", var))
  logit = glm(form, data = PIAAC_clean, family = "binomial")
  OR = data.frame(exp(cbind(OR = coef(logit), confint(logit))))[-1, ]
  OR = cbind(OR, coef(summary(logit))[-1, 4])
  colnames(OR) <- c("OR", "2.5%", "97.5%", "p-val")
  return(OR)
}

varlist <- c("numeracy", "educ", "income.cat", "race", "literacy", "problem_solving")
univ_summary <- do.call(rbind.data.frame, lapply(varlist, univ))
univ_summary$sig <- ifelse(univ_summary$`p-val` < 0.05, "*", "")
kable(univ_summary)%>%
  kable_styling(full_width = F) 
```

**With health status 2 as outcome**    
health_status_2:     
  Not Good: health score >=4;    
  Good: health score < 4.   
  
```{r}
univ <- function(var){
  form = formula(paste0("health_status_2 ~ ", var))
  logit = glm(form, data = PIAAC_clean, family = "binomial")
  OR = data.frame(exp(cbind(OR = coef(logit), confint(logit))))[-1, ]
  OR = cbind(OR, coef(summary(logit))[-1, 4])
  colnames(OR) <- c("OR", "2.5%", "97.5%", "p-val")
  return(OR)
}

varlist <- c("numeracy", "educ", "income.cat", "race", "literacy", "problem_solving")
univ_summary <- do.call(rbind.data.frame, lapply(varlist, univ))
univ_summary$sig <- ifelse(univ_summary$`p-val` < 0.05, "*", "")
kable(univ_summary)%>%
  kable_styling(full_width = F) 
```

The above two univariate results show that all these covariates are significant predictors.   

## Step-wise multivariate logistic regression     
**With health status 1 as outcome**    
Outcome: health_status_1         
  Poor: health score >=5;    
  Not poor: health score < 5. 
Predictor: numeracy score        
Covariates: education, income and race.    
```{r, warning=FALSE}
full1 <- glm(health_status_1 ~ numeracy + educ + income.cat + race, data = PIAAC_clean, 
            family = "binomial")
final1 <- step(full1, direction = "both", trace = 0)
multi_summary1 <- exp(cbind(coef(final1), confint(final1))[-1, ]) %>% as.data.frame()
colnames(multi_summary1)[1] <- "OR"
multi_summary1$sig <- ifelse(summary(final1)$coefficients[-1, 4] < 0.05, "*", "")
kable(multi_summary1) 
```

```{r, message=FALSE}
# ROC curve
rocplot <- function(truth, pred, ...){
  predob = prediction(pred, truth)
  perf = performance(predob, "tpr", "fpr")
  plot(perf, colorize=F, ...) 
  area = auc(truth, pred)
  area = format(round(area, 4), nsmall = 4)
  text(x = .8, y = .1, labels = paste("AUC = ", area))
  segments(x0 = 0, y0 = 0, x1 = 1, y1 = 1, col = "gray", lty = 2)
}

rocplot(PIAAC_clean$health_status_1, predict.glm(final1, PIAAC_clean, type = "response", na.action = na.omit), title = "ROC plot")
title(main = "ROC plot")
```


**With health status 2 as outcome**    
Outcome: health_status_1         
  Not Good: health score >=4;    
  Good: health score < 4. 
Predictor: numeracy score        
Covariates: education, income and race.     

```{r, warning=FALSE}
full2 <- glm(health_status_2 ~ numeracy + educ + income.cat + race, 
             data = PIAAC_clean, family = "binomial")
final2 <- step(full2, direction = "both", trace = 0)
multi_summary2 <- exp(cbind(coef(final2), confint(final2))[-1, ]) %>%
  as.data.frame()
colnames(multi_summary2)[1] <- "OR"
multi_summary2$sig <- ifelse(summary(final2)$coefficients[-1, 4] < 0.05, "*", "")
kable(multi_summary2) 
```

```{r}
rocplot(PIAAC_clean$health_status_2, predict.glm(final2, PIAAC_clean, type = "response", na.action = na.omit), title = "ROC plot")
title(main = "ROC plot")

```

## Considering the additional predictors: Literacy, Problem solving abilities.    

Outcome: health_status_1         
  Poor: health score >=5;    
  Not Poor: health score < 4. 
Predictor: numeracy score        
Covariates: education, income, race, literacy and problem solving.     


```{r}
PIAAC_clean2 <- na.omit(PIAAC_clean) 

```


```{r, warning=FALSE}
full3 <- glm(health_status_1 ~ numeracy + educ + income.cat + race +
               literacy + problem_solving, data = PIAAC_clean, 
             family = "binomial")
final3 <- step(full3, direction = "both", trace = 0)
multi_summary3 <- exp(cbind(coef(final3), confint(final3))[-1, ]) %>%
  as.data.frame()
colnames(multi_summary3)[1] <- "OR"
multi_summary3$sig <- ifelse(summary(final3)$coefficients[-1, 4] < 0.05, "*", "")
kable(multi_summary3)
```

```{r}
rocplot(PIAAC_clean2$health_status_1, predict.glm(final3, PIAAC_clean2, type = "response", na.action = na.omit), title = "ROC plot")
title(main = "ROC plot")

```

```{r, warning=FALSE}
full4 <- glm(health_status_2 ~ numeracy + educ + income.cat + race +
               literacy + problem_solving, 
             data = PIAAC_clean, family = "binomial")
final4 <- step(full4, direction = "both", trace = 0)
multi_summary4 <- exp(cbind(coef(final4), confint(final4))[-1, ]) %>%
  as.data.frame()
colnames(multi_summary4)[1] <- "OR"
multi_summary4$sig <- ifelse(summary(final4)$coefficients[-1, 4] < 0.05, "*", "")
kable(multi_summary4)
```

```{r}
rocplot(PIAAC_clean2$health_status_2, predict.glm(final4, PIAAC_clean2, type = "response", na.action = na.omit), title = "ROC plot")
title(main = "ROC plot")
```
