---
title: "RVq2"
author: "Haojia Li"
date: "6/6/2020"
output: 
  html_document:
    code_folding: hide
    theme: readable
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(scipen = 3, digits = 3)
```

```{r packages}
# data manipulation
library(tidyverse)
# knit tables
library(knitr)
library(kableExtra)
```

```{r load data}
load("./data/ESP.RData")
```

## RVq2: care provider visits in the past 12 months

### Distribution (barplot)

```{r}
# table(ESP$visits_12mos, useNA = "ifany")
# median(ESP$RVq2)
ESP %>%
  ggplot() +
  # draw the basic bar plot
  geom_bar(aes(x = visits_12mos), stat = "count", fill = "white", color = "black") +
  # add the count label to each bar
  geom_text(stat = "count", aes(x = visits_12mos, label = ..count..), position = position_stack(vjust = 0.5), size = 3.2) +
  # add a vertical dash line showing the position of median visit level
  # the median level is 3: once every few month
  # the first half of the sample is made up of:
    # 72 never + 237 1-2 times + 91 once every few month
    # 91 is about 1/3 of the whole "once every few month" group
  # thus the position of the dash line is at about 1/3 of the bar of "once every few month
  geom_vline(xintercept = 2.85,linetype = "longdash", color = "red", ) +
  theme_bw() +
  # rotate the x-axis labels to avoid overlap
  theme(axis.text.x = element_text(angle = 20, hjust = 1)) +
  labs(x = NULL, y = NULL, 
       title = "Distribution of care provider visits in the past 12 months", 
       subtitle = 'The median level of visits is "once every few month"\nDash line: the approximate position of the median in the sample')
```

## RVq2 and JAq4

### Distribution

```{r}
ESP %>%
  group_by(visits_12mos, health_status) %>%
  # summarise the number in each combination of care provider visits and health status
  summarise(N = n()) %>%
  # fill each tile by N
  # no matter what color I set here, the output plot always shows the same color, which is weird
  ggplot(aes(x=visits_12mos, y=health_status, fill=N, color = "")) +
  # plot heatmap
  geom_tile() + 
  # add the count label to each tile
  geom_text(aes(label = N), size = 4) + 
  # change direction to reverse the order of the color corresponding to the count levels
  scale_fill_distiller(palette = "Blues", direction = "horizontal") +
  theme_minimal() +
  # rotate the x-axis labels to avoid overlap
  theme(axis.text.x = element_text(angle = 20, hjust = 1)) +
  labs(x = NULL, y = NULL, title = "Distribution of participants by health status and care provider visits") +
  # omit the legend of color
  guides(color = F)
```

### Correlation: Cramer's V

Since both of RVq2 (care provider visits) and JAq4 (health_status) are categorical variables, traditional test such as chi-square test and fisher's test is not applicable. We calculate Cramer's V based on chi-square statistic here to evaluate the correlation between the two outcome of interest. **The smaller V, the lower the correlation.**

[Overview of correlation measures between categorical and continuous variables](https://medium.com/@outside2SDs/an-overview-of-correlation-measures-between-categorical-and-continuous-variables-4c7f85610365)

[Wikipedia of Cramer's V](https://en.wikipedia.org/wiki/Cram%C3%A9r%27s_V)

[Implementation in R](https://datascience.stackexchange.com/questions/893/how-to-get-correlation-between-two-categorical-variable-and-a-categorical-variab)

```{r}
# v.cclt (Cramer's V Calculator) is a function calculating Cramer's V
v.cclt <- function(tbl) {
  chi2 <- chisq.test(tbl, correct = F)
  v <- sqrt(chi2$statistic / sum(tbl))
  names(v) <- NULL
  return(v)
}

tbl <- table(ESP$visits_12mos, ESP$health_status)
v.cclt(tbl)
```

As V = 0.293, there is no strong correlation between RVq2 (care provider visits) and JAq4 (health_status).

## RVq2 and numeracy score

```{r}
ESP %>%
  ggplot() +
  geom_boxplot(aes(x=visits_12mos, y=numeracy)) +
  theme_bw() +
  # rotate the x-axis labels to avoid overlap
  theme(axis.text.x = element_text(angle = 20, hjust = 1)) +
  labs(x = NULL, y = "Numeracy Score", title = "Boxplot of numeracy score by care provider visits")

# performe one-way ANOVA test
summary(aov(numeracy~visits_12mos, data = ESP))
```

There is no significant difference across the RVq2 groups regarding numeracy score.

## RVq2 and covariates

### Correlation: Cramer's V

```{r}
covariates <- c("gender", "age.cat", "educ", "income.cat", "race", "hisp")
v.values <- sapply(ESP[covariates],function(x) v.cclt(table(ESP$visits_12mos,x)))
v.values <- t(v.values) %>% as.data.frame()
rownames(v.values) <- "V"
kable(v.values) %>%
  kable_styling(full_width = F)
```
