---
title: "ESP - Data Analysis"
author: 
- "Haojia Li"
- "Tianran Zhang"
- "Zihan Yang"
date: "5/29/2020"
output: 
  html_document:
    code_folding: hide
    theme: readable
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(scipen = 3, digits = 3)
# set working directory into the Data folder
# Data folder including 2 sub-folder: 
# 1.  RawData(xlsx, csv, sav) 2. RData(RData)
# path <- "../Data"
# knitr::opts_knit$set(root.dir = path)
```

```{r packages}
# read data
library(haven)
library(readxl)

# data manipulation
library(tidyverse)
# knit tables
library(knitr)
library(kableExtra)

# install.packages("RVAideMemoire")
library(RVAideMemoire) # fisher.multcomp with fdr
# install.packages("rstatix")
library(rstatix) # fisher_test with fdr

# Regression on weighted data
library(survey)
# PIAAC Regression
library(intsvy)
```



```{r load data}
# Load data from the Rdata folder in the Data folder
load("../Data/ESP.RData")
load("../Data/NYS.RData")
load("../Data/PIAAC.RData")
```

# ESP (no weighted)

**Empire State Poll (ESP)**

Step-wise multivariate linear regression

* Outcome: binary health status    
1. Original Levels (Multi-level/Excellent/Very good/Good/Fair/Poor/Unknown)

2. Health status 1(Binary): Not Poor VS. Poor

3. Health status 2 (Binary): Good VS. Not Good

* Predictor: numeracy score    
* Covariates: education, ethincity, income, race    
* Interaction term: education & numeracy score    


| Outcome                | Health status 1 (Not Poor VS. Poor) | Health status 2 (Good VS. Not Good) |
| ---------------------- | ----------------------------------------------- | ----------------------------------------------- |
| significant predictor  | numeracy score                                            |                                                 |
| significant covariates | Education, ethinicity                                     | income, education                               |
| AUC                    | 0.8123                                                    | 0.7024                                          |

**Program for the International Assessment of Adult Competencies (PIAAC 2017)**     

Step-wise multivariate linear regression       

* Outcome: binary health status  
* predictor: numeracy score    
* covariates: income, race, education    
* Interaction term: education & numeracy score    

| Outcome                | Health status 1 (Not Poor VS. Poor) | Health status 2 (Good VS. Not Good) |
| ---------------------- | ----------------------------------------------- | ----------------------------------------------- |
| significant predictor  | numeracy score                                  | numeracy score                                  |
| significant covariates | income, race                                    | income, race, education                         |
| AUC                    | 0.7524                                          | 0.7114                                          |



Considering additional covariates:

* Outcome: binary health status     
* predictor: numeracy score, literacy performance, problem solving      
* covariates: income, race, education      

| Outcome                | Health status 1 (Not Poor VS. Poor) | Health status 2 (Good VS. Not Good) |
| ---------------------- | ----------------------------------------------- | ----------------------------------------------- |
| significant predictor  | numeracy score                                  | numeracy score                                  |
| significant covariates | income, race, education, problem solving        | income, race, education, problem solving        |
| AUC                    | 0.8196                                          | 0.6920                                          |


1. `numeracy score` is a significant predictor, `income` and `race` are significant covariates in all models;
2. `education` is a significant covariates in all the other models except for the regression of PIAAC health_status_1 (Poor VS. Not Poor) with 4 independent variables (no additional predictors).     
3. For additional predictors, `problem solving` is significant while `literacy` is not significant no matter which cut point to set.     

## Refusal Rate
```{r refuse}
# calculate the number of refuse or -1 (in JAqs) in each column
refuse <- sapply(ESP, function(x) sum(x%in%c("Refused",-1)))
# keep the columns that have refuse
refuse <- refuse[which(refuse>0)]
# append the refuse rate and create a data frame; -2 means remove `age` since it overlaps with age.cat
refuse.df <- rbind(refuse, refuse/8) %>% as.data.frame()
refuse.df <- refuse.df[,-2]
# set col- and rownames
colnames(refuse.df) <- c("Gender", "Age Level", "Education", "Race", "Ethnicity", "JAq1: Fractions", "JAq2: Percentages", "JAq3: Numerical Info")
rownames(refuse.df) <- c("Number", "%")
kable(refuse.df) %>% kable_styling()
```

## Comparing means of numeracy score by health status
Relationship between numeracy score and health status
```{r}
par(mfrow = c(1,2))
plot(x=ESP$numeracy, y=ESP$JAq4, pch=19, xlab = "Numeracy Score", ylab = "Health Status Score", main = "Scatter plot of numeracy score and health status")
# boxplot(numeracy~JAq4, data = ESP, xlab = "Health Status Score", ylab = "Numeracy Score", main = "Box plot of numeracy score and health status")

boxplot(numeracy~JAq4, data = ESP, xlab = "Health Status Score", ylab = "Numeracy Score", main = "Box plot of numeracy score and health status")

ESP_plot <- ESP%>%
  mutate(health_1 = case_when(
    JAq4%in%c(1:3) ~ "Good",
    JAq4%in%c(4:5) ~ "Not Good",
    ))%>%
  mutate(health_2 = case_when(
    JAq4%in%c(1:4) ~ "Not Poor",
    JAq4%in%c(5) ~ "Poor",
    ))%>%
  select(numeracy,health_1,health_2)

ggplot(ESP_plot, aes(x=health_1, y=numeracy, fill=health_1)) + 
  geom_boxplot()+
  labs(title="Box plot of numeracy score and health status 1  (Not Poor VS. Poor) ",y="Numeracy Score", x = "Health Status Score")+theme_bw()+
  theme(plot.title = element_text(hjust = 0.5,size = 18))+scale_fill_brewer(palette="Dark2")

ggplot(ESP_plot, aes(x=health_2, y=numeracy, fill=health_2)) + 
  geom_boxplot()+
  labs(title="Box plot of numeracy score and health status 2  (Good VS.Not Good) ",y="Numeracy Score", x = "Health Status Score")+theme_bw()+
  theme(plot.title = element_text(hjust = 0.5,size = 18))+scale_fill_brewer(palette="Dark2")


# perform one-way ANOVA test
summary(aov(numeracy~health_status, data = ESP))
```

## Comparing means of numeracy score by education
Relationship between numeracy score and education

```{r}
ESP %>%
  filter(educ != "Refused") %>%
  group_by(educ) %>%
  ggplot() +
  geom_boxplot(aes(x=educ, y=numeracy)) +
  labs(x = NULL, title = "Box plot of numeracy score and education") 
  

# performe one-way ANOVA test
summary(aov(numeracy~educ, data = ESP))
```

# ESP vs. NYS (Fisher's test)

```{r}
cat.smry <- function(mat, labels = NULL) {
  perc <- paste0(round(prop.table(mat)*100,2),"%")
  n.perc <- paste0(mat,"(",perc,")")
  if(is.null(labels))
    names(n.perc) <- names(mat)
  else
    names(n.perc) <- labels
  return(n.perc)
}
# gather all the cat.smry result for ESP sample
ESP.smry <- c(cat.smry(table(ESP$gender)),
              cat.smry(table(ESP$age.cat)),
              cat.smry(table(ESP$educ)),
              cat.smry(table(ESP$income.cat)),
              cat.smry(table(ESP$race)),
              cat.smry(table(ESP$hisp)))

# join NYS.smry and ESP.smry together (drop levels such as "No data" and "Refused" which are unique for ESP sample by left join NYS)
comp1 <- data.frame(Demographics = rownames(NYS), NYS = NYS$n_perc) %>%
  left_join(data.frame(d = names(ESP.smry), ESP = ESP.smry), by = c("Demographics" = "d"))

# extract counts from ESP.smry
ESP.smry <- comp1$ESP
ESP.n <- c()
for(i in 1:length(ESP.smry)) {
  n.perc <- ESP.smry[i]
  n <- substr(n.perc, 1, unlist(gregexpr("\\(",n.perc))-1)
  ESP.n <- c(ESP.n, n)
}
ESP.n <- as.numeric(ESP.n)
# bind the counts of ESP and NYS preparing for fisher's test
fisher.mat <- rbind(ESP.n, unlist(NYS$counts)) %>% as.matrix()


# add subtitles of demographics
comp1 <- rbind(c("Gender", "", ""), comp1[1:2,],
              c("Age", "", ""), comp1[3:6,],
              c("Education", "", ""), comp1[7:9,],
              c("Income", "", ""), comp1[10:14,],
              c("Race", "", ""), comp1[15:19,],
              c("Ethnicity", "", ""), comp1[20:21,]) %>% as.data.frame()

# append p-values of fisher's test
comp1$pval <- sapply(list(1:2,3:6,7:9,10:14,15:19,20:21), 
                     function(x){
                       mat <- fisher.mat[,x]
                       # pval <- fisher_test(mat)$p.value
                       # pval <- fisher.multcomp(mat)$p.signif
                       pval <- fisher.test(mat, simulate.p.value = T)$p.value
                       pval <- ifelse(pval<0.001, "<0.001", round(pval,3))
                       c(pval, rep("",length(x)))
                       }) %>% unlist()

kable(comp1, row.names = F) %>% 
  kable_styling(full_width = F) %>%
  row_spec(c(1,4,9,13,19,25), bold = T)

```

The first two levels of household income for NYS population are actually <$15k and $15k-50k.


*Question: Do you have any suggestions about the method of how to split people with household income <$15k to <$10k and $10k-15k?*

(Comment: Less level with 0-50, 50-100, 100+) <br>
(Comment: Do we need consider post-hoc analysis for multiple pariwise fisher's test. I search online the answer is we could use the same method like post-hoc for anova in exact fisher's test. Then I tried to use adjusted p-values in fisher's test here with these two functions in the above table, but not working) <br>
[fisher_test](https://rpkgs.datanovia.com/rstatix/reference/fisher_test.html) <br>
[fisher.multcomp](https://rdrr.io/cran/RVAideMemoire/man/fisher.multcomp.html) <br>

## ESP (with weights)
Based on the summary table above, education, income and race are three variables that have different different distributions in ESP data and NYS dataset. However, among these three variables only education is a significant covariate in previous regression results. It appears the distributions of income and race are not associated with people's health status. So, here I only add survey weights to adjust the educationl level distribution in ESP data.    

*Question: Do we also need to add survey weights on income and race even though they are not significant covariates?* 
(Commnet:<br>
1. adjusted p-value mignt change a little bit from the unadjusted one.<br>
2. The levels of income would change from the current one (balance or power) <br>
3. For observation purporse, we might consider add income and race, the weight value would change 
4. The weight would change if we will combine the design bias. <br>
Therefore, the one given aboveis the optimized solution we could provide with available information now.br>


*Question: What is the final supposed goal for here PIACC vs. ESP and ESP vs. NYS： Inference /Observation/Prediction?*

*Question: So far we only consider the post-strafication weights without considering the design weights considering the sampling ratio.*
(EG: Table 3 in [ESP methodology](https://sri.cornell.edu/sri/polls/files/esp/2019/ESP19%20-%20Introduction%20and%20Methodology%20Report.pdf))

## Info: Weights

1. Weights are common used in sample survey for adjusting: <br>

a. Sample design (oversampling of some groups or disproportionate stratification)<br>

b. Nonresponse<br>

c. Bring sample backto being representative of the population ona select set of demographic characteristics <br>

2. Cnsensus on useage of weight: <br>
a. General consensus in the literature that weights should be use for descriptive statistics (e.g., Kish & Frankel, 1974) 
b. Less consensus on whether weights should be routinely used in multivariate
models, such as regression (e.g., Gelman, 2007 and comments; Kott, 2007; Winship & Radbill, 1994)

#### Pros
a. Allow us to claim our results are representative of the population (at least that they have the same distribution on major demographic characteristics as the population)

b. Used to adjust descriptive information on the sample 

#### Cons
They introduce a substantial design effect into our data <br>

a. Weights will increase the standard errors of our statistics making our findings less precise and more variable <br>

b. The larger the variability in the weights the larger the design effects<br>

c. This will affect most kinds of statistical analysis <br>
–  Descriptive statistics (means, percentages)
–  Regression coefficients
–  Other multivariate coefficients

3. Method to test if the weight is needed: <br>

a. Sensitivity analysis—compare the coefficients of interest from analysis with and without weights. If not significantly different then weights unnecessary.<br>

b. Add the weight and the interaction of the weight with each independent variable to the model. If these do not add significant amount of explained variance to the model, then weights not necessary.<br>

c. When the weights have no effect on the parameter
estimates they are called ignorable.<br>

4. Recommendation <br>

a.  Compare basic substantive model with and without weights and including model-based variables. If same coefficients, unweighted data would be best because of smaller standard errors. <br>

b. Test if a model with the weight and the interactions of the weight with all independent variables included adds significant explained variance. If yes, then use the weighted data or use a model-based approach.<br>

5. Types of weights
a. Design Weights: related to sampling fraction(EG: Table 3 in [ESP methodology](https://sri.cornell.edu/sri/polls/files/esp/2019/ESP19%20-%20Introduction%20and%20Methodology%20Report.pdf))
* Question here*
b. Post-Stratification weights: realted to the demographical freatures of population 
Compute weights separately but sequentially.

**Using manual iterative solution**

```{r}
ESP.weighted <- ESP %>%
  filter(hisp != "Refused" & educ != "Refused") %>%
  mutate(hisp = factor(hisp), educ = factor(educ), wt = 1)


# This function generate survey weights automatically
surv.wt <- function(x, w, col_pos, col_name){
  # x: dataset of interest
  # w: NYS percentages
  # cpl_pos: column number of the variable to be weighted
  # col_name: name of the variable to be weighted
  
  x1 <- x %>%
    group_by_at(col_pos) %>%
    summarise(wt = sum(wt)/nrow(x)) %>%
    mutate(wt_new = w/wt) %>%
    select(-wt)
  
  merge(x, x1, by = col_name) %>%
    mutate(wt = wt * wt_new) %>%
    select(-wt_new)
}

```

```{r}
# Manully input the NYS education level proportion and hispanic proportion

wt.edu <- c(0.3892, 0.2777, 0.3332)
# wt.hisp <- c(0.1647, 0.8353)


ESP.weighted <- surv.wt(ESP.weighted, wt.edu, 4, "educ")
# ESP.weighted <- surv.wt(ESP.weighted, wt.hisp, c1 = 7, c2 = "hisp")

```



## Regression on weighted data        

**Regression with health status 1 as the outcome**   

```{r}
ESP.design <-svydesign(
		id = ~0,
		data = ESP.weighted ,
		weights = ~ wt
	)

sum1 <- summary(mysvyglm <- svyglm(health_status_1 ~ hisp + educ + numeracy, ESP.design,
                           family=quasibinomial()))
sum1 <- data.frame(sum1$coefficients[-1, c(1, 4)])
colnames(sum1) <- c("Estimate", "P-value")
sum1$sig <- ifelse(sum1$`P-value` <0.05, "*", "")


kable(sum1) %>%
  kable_styling(full_width = F) 
```

**Regression with health status 2 as the outcome**     

```{r}
sum2 <- summary(mysvyglm <- svyglm(health_status_2 ~ hisp + educ + numeracy, ESP.design,
                           family=quasibinomial()))
sum2 <- data.frame(sum2$coefficients[-1, c(1, 4)])
colnames(sum2) <- c("Estimate", "P-value")
sum2$sig <- ifelse(sum2$`P-value` <0.05, "*", "")


kable(sum2) %>%
  kable_styling(full_width = F) 

```

**Interesting finding**  
In the first model with health status 1 as the outcome, educational level and numeracy are significant. While in the second model with health status 2 as the outcome, ethnicity and numeracy are significant.    


# ESP vs. PIAAC (education, income and ethnicity)

Race is not comparable between the two samples since PIAAC combines race and ethnicity together as one variable.   

- Solution: create a new variable named "hisp" indicated whether one person is hispanic or not.   

The levels of the B_Q01A variable in PIAAC dataset matched the content of ISCED-97.

```{r}
PIAAC <- PIAAC %>%
  mutate(hisp = case_when(
    race == "Hispanic" ~ "Hispanic",
    race %in% c("White", "Black", "Asian/pacific islander", "Other") ~ 
      "Npn-Hispanic",
    race == "Unknown" ~ "Unknown"))
```

```{r}
# bind ESP and PIAAC together
demo <- rbind(data.frame(Study = "ESP", Education = ESP$educ, 
                         Income = ESP$income.cat, Hisp = ESP$hisp),
              data.frame(Study = "PIAAC", Education = PIAAC$educ, 
                         Income = PIAAC$income.cat, Hisp = PIAAC$hisp)) %>% as.data.frame()

demo <- demo %>%
  mutate(Education = factor(Education), Income = factor(Income), 
         Hisp = factor(Hisp))

# gather cat.smry results for ESP and PIAAC
## -4 here means drop the "Refused" level in ESP$educ and the "Unknown" level in PIAAC$educ since each of them is the unique level for its study
## if we keep these two levels, we will find that there is no p-value output of compareGroup function
ESP.smry1 <- c(cat.smry(table(ESP$educ))[-4],cat.smry(table(ESP$income.cat)), cat.smry(table(ESP$hisp)))

PIAAC.smry <- c(cat.smry(table(PIAAC$educ))[-4],cat.smry(table(PIAAC$income.cat)), cat.smry(table(PIAAC$hisp)))

comp2 <- data.frame(Demographics=names(ESP.smry1), ESP=ESP.smry1, PIAAC=PIAAC.smry)
# add subtitles of demographics
comp2 <- rbind(c("Education", "", ""), comp2[1:3,],
               c("Income", "", ""), comp2[4:9,],
               c("Hisp", "", ""), comp2[10:12, ]) %>% as.data.frame()
rownames(comp2) <- NULL

# perform Fisher's test and append p-values to the data frame
p.educ <- fisher.test(table(demo$Study, demo$Education), simulate.p.value = T)$p.value
p.educ <- ifelse(p.educ<0.001, "0.001", round(p.educ,3))

p.income <- fisher.test(table(demo$Study, demo$Income), simulate.p.value = T)$p.value
p.income <- ifelse(p.income<0.001, "<0.001", round(p.income, 3)) 


p.hisp <- fisher.test(table(demo$Study, demo$Hisp), simulate.p.value = T)$p.value
p.hisp <- ifelse(p.hisp<0.001, "<0.001", round(p.hisp, 3)) 
comp2$pval <- c(p.educ, rep("",3), p.income, rep("",6), p.hisp, rep("", 3))

kable(comp2, row.names = F) %>%
  kable_styling(full_width = F) %>%
  row_spec(c(1,5, 12), bold = T)
```


## Info: Plausible value in PIAAC

PIAAC cannot provide individual-level results, because each adult answers only a small number of assessment questions. PIAAC provides reliable estimates of proficiency only at the national level, or at the level of large subgroups (e.g., females, employed, or college educated).       

Plausible Values (PVs) allows PIAAC dataset information to be saved at the case level to estimate proficiency at the national or subgroup level. Each case's PVs reflect not only that individual's performance on the small number of items s/he answered, but also the performance of similar respondents on the rest of the PIAAC assessment.    


Each individual case in the PIAAC dataset has a randomly chosen set of ten plausible values (PVs). All ten PVs must be used together to estimate proficiency, or else one understates the variability in the predicted outcomes. The randomly chosen set of PVs best represents the score distribution for a subgroup of adults.    

reference:     
https://cran.r-project.org/web/packages/intsvy/intsvy.pdf    
https://nces.ed.gov/training/datauser/PIAAC_04/assets/PIAAC_04.pdf    
https://www.oecd.org/skills/piaac/piactools_16oct_for_web.pdf    

(Comment: True, same as what I fould last week, therefore I was wondering why we need to pick the PV1 specifically.)

```{r}
raw_PIAAC <- read_spss("../Data/prgusap1_puf.sav")
num_pv <- paste0("PVNUM", 1:10)
# paste(num_pv, collapse = " + ")

pv <- raw_PIAAC[, num_pv] %>%
  gather(key = "PVs", value = "value")
#table1(~ PVNUM1 + PVNUM2 + PVNUM3 + PVNUM4 + PVNUM5 + PVNUM6 + 
#         PVNUM7 + PVNUM8 + PVNUM9 + PVNUM10, data = pv)

ggplot(data = pv) +
  geom_boxplot(aes(x = PVs, y = value)) + 
  scale_x_discrete() +
  ggtitle("Boxplots among 10 numeracy PVs") +
  ylab("PV score")

summary(aov(value ~ PVs, data = pv)) 
```

ANOVA test shows that there is no significant difference between these 10 PVs. 


## Correlation: health status vs. numeracy score 
The numeracy proficiency score are calculated with the implemented function "piaac.mean.pv" in the package "intsvy". This function uses ten plausible values to calculate the mean achievement score and its standard error.    

The t-test statistics are calculated using the online t-test calculator:  
https://select-statistics.co.uk/calculators/two-sample-t-test-calculator/

Health Staus 1 : Not Poor VS. Poor <br>
Health status 2: Good VS. Not Good <br>



```{r}
PIAAC.pv <- cbind(PIAAC[, c("educ", "income.cat", "race",
                            "health_status_1", "health_status_2")],
                  raw_PIAAC)

pv1 <- piaac.mean.pv(pvlabel = "NUM", by = "health_status_1", data = PIAAC.pv)[-3, 1:4]
pv1$p_val <- c("<0.001", "")
pv1 %>%
  kable()%>%
  kable_styling(full_width = F)

pv2 <- piaac.mean.pv(pvlabel = "NUM", by = "health_status_2", data = PIAAC.pv) [-3, 1:4]
pv2$p_val <- c("<0.001", "")
pv2 %>%
  kable()%>%
  kable_styling(full_width = F)

```


Health Staus 1 : Not Poor VS. Poor <br>
Health status 2: Good VS. Not Good <br>

The two-sample t-test results show that numeracy score is positively associated with people's health status.   










